\section{Introduction}

Our goal is to understand and articulate the characteristics of
cloud-based dataflow processing in the context of HPC analysis tasks,
implement sample apps using spark, pegasus, and mpi and evaluate their
performance on both a supercomputing cluster and a cloud.

We will compare XYZ written in Apache Spark vs. MPI (a traditionally
used messaging middleware)

The microbenchmarks are chosen based on typical
computation/communication patterns found in HPC.

We're also looking at Python vs. Scala and whether there are any
performance differences.

Some words about tuning garbage collection. These folks seem to have
done some good legwork on same\footnote{https://databricks.com/blog/2015/05/28/tuning-java-garbage-collection-for-spark-applications.html}.

Also simply describe differences between implementation (e.g. lines of
code).

Out of core: Implicit in Spark

NOTE: we need to clarify what we mean by \emph{dataflow.} Notably, it is a different from its meaning in \emph{compilers} and \emph{computer architecture}:
\begin{itemize}
\item
  \href{https://en.wikipedia.org/wiki/Data-flow_analysis}{\emph{https://en.wikipedia.org/wiki/Data-flow\_analysis}}
\item
  https://en.wikipedia.org/wiki/Dataflow\_architecture
\end{itemize}

\paragraph{Project Resources}\label{project-resources}

\textbf{Github:}
\href{http://github.com/cchriste/dataflow}{\emph{http://github.com/cchriste/dataflow}}

\textbf{Slack:}
\href{https://dataflowanalysis.slack.com}{\emph{https://dataflowanalysis.slack.com}}

\textbf{Google Docs:}
\href{https://drive.google.com/folderview?id=0BxLkEMNd9q6FfmpRaFZXSGlPc0JsSDdVdndCUm83SzN6UnlLVEk5T3ZsZmJ0VEVGREtNTkE\&usp=sharing}{\emph{dataflow\_analysis}}

\textbf{Dropbox:}
\href{https://www.dropbox.com/sh/odsd9uxe1elbhbf/AAD8J1TGuFY1VHJl2oPdm5E0a?dl=0}{\emph{spark\_hpc}}


